---
title       : HR Analysis
subtitle    : Predicting the probability of a candidate joining a Company
author      : Nithya
job         : Data Science
framework   : revealjs      # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

```{r echo=FALSE}

library(knitr)
library(ggplot2)
library(reshape2)

```

### Problem Statement Overview

The HR executives of a company continuously face problems due to uncertainly associated with a Candidate Joining. 

&nbsp;

#### Challenges faced by HRs
- Same exhaustive recruitment process for all candidates irrespective of they join / not join
- Do not have sufficient information to prioritize the candidates
- Too many Offered Candidates to concentrate on within limited time

&nbsp;

#### Challenges faced by Companies -
- Oversized HR teams
- High HR budget spend
- HR Department is always a cost to a company with no ROI
- Poor productivity of HR team due to Poor planning of recruitment activities

--- .class #id 

### Our Analysis and Solution Overview -

Our model predicts the probability of a Candidate Joining based on inputs collected by the HR from the applicants and help the HR consultants in making their recruitment efforts more productive. In turn, saving a lot of dollars for the company. We are using a Generalized Logistic Regression Model 

```{r echo=FALSE}

url2 <- "https://www.dropbox.com/s/oq973dlyr0qp4cn/tab_data.csv?dl=1"
download.file(url2, "tab_data.csv", quiet = TRUE)
          
url3 <- "https://www.dropbox.com/s/vbt9b8ypgy6819b/fit_data.csv?dl=1"
download.file(url3, "fit_data.csv", quiet = TRUE)

tab <- read.csv("tab_data.csv", header = TRUE)
fit_data <- read.csv("fit_data.csv", header = TRUE)
fit_data[, 2] <- round(fit_data[, 2],2)

```

&nbsp;

Here are the model performance parameters

```{r echo=FALSE}
d <- fit_data
rownames(d) <- d[,1]
d <- data.frame(t(d))
d <- d[2,]
kable(d[, c(1,2,3,4,6)], row.names = FALSE)
```

&nbsp;

The accuracy of our model is : `r round((100 * fit_data[1,2]),2)` %. This is the best we could get given the really poor correlation between the parameters and outcome.

--- .class #id 

### Model Analysis

Below is the results of a sample test done on 12000 (approx.) observations. As we can see, the proportion of False-Positives and False-Negatives is very less and the model is able to predict the outcome to a fairly reasonable extent.
```{r echo=FALSE}

ggplot(tab, aes(x = actual, y = predicted, color = as.factor(difference))) + geom_jitter() + scale_colour_manual(values=c("darkgreen", "coral"),  name = "Joined ?",labels=c("1 = Yes", "0 = No"))

```

--- .class #id 

### Tool Introduction

We also have created a Shiny App tool to implement the predection model. This handy tool provides:

&nbsp;

#### For consumption of HR Consultants - 
- Easy interface to the HR consultants 
- Prediction of the probability of a candidate joining

&nbsp;

#### For consumption of Analysts - 
- Key information about the model
- Reference to the data analysis during modeling

&nbsp;

#### Important URLs - 
- Shiny App Link -> https://nithsubr.shinyapps.io/HR_Analysis
- Supporting files -> https://github.com/nithsubr/HR_Analysis

--- .class #id 

### Tool Help

```{r fig.width=10, fig.height=10, echo=FALSE}
library(png)
library(grid)
img <- readPNG("Tutor.png")
 grid.raster(img)
```
